---
title: "SBL Survey Analysis"
output:
  html_document:
    toc: yes
---

------------------------------------------------------------------------

# Supplementary Material Chapter 4

**Author:** Michael S. Feurstein\
**Created:** 15 May 2025\
**Last compiled:** `r format(Sys.time(), '%d %B %Y')`

------------------------------------------------------------------------

## About this notebook

This notebook is a supplementary material for Chapter "Perceived Usability Survey" used for replication of results. Data preparation steps have been done before. All further details on survey design, methodology and analysis procedure can be found in the corresponding Chapter of thesis.

------------------------------------------------------------------------

### Libraries

```{r message=FALSE, warning=FALSE}
library(dplyr)
if(!require(likert)){install.packages("likert")}
library(likert)
if(!require(svglite)){install.packages("svglite")}
library(svglite)
library(ggplot2)
library(ggtext)
if(!require(lemon)){install.packages("lemon")}
library(lemon)
library(stringr)
if(!require(ggpubr)){install.packages("ggpubr")}
library(ggpubr)
```

### Load Data

File: ***"data_survey_prepared.csv"*** - contains calculated sus-score.
```{r}
# read from prepared csv
df <- read.csv("data_prepared/data-survey-sus_prepared.csv")
df <- subset(df, select = -c(X))
```

File: ***participant_profiles_anon_prepared.csv*** - contains all anonymized data on participants and their profile (self assessed role: author, user, researcher, other and meta data on activity in field: no of videos, no of papers, no of courses etc.)
```{r}
# read from prepared csv
profiledata <- read.csv("data_prepared/data-survey-profiles_prepared.csv")
```

Files: ***sus_sc_likert_tidy.csv***, ***sus_cnl_likert_tidy.csv*** and ***sus_cnl_likert_tidy.csv*** - contain individual responses to the System Usability Scale questions.
```{r}
df_sc_tidy <- read.csv("data_prepared/sus_sc_likert_tidy.csv")
df_cnl_tidy <- read.csv("data_prepared/sus_cnl_likert_tidy.csv")
df_kv_tidy <- read.csv("data_prepared/sus_kv_likert_tidy.csv")
```

File: ***data-survey-ranking_prepared.csv*** contains ranking 1st, 2nd, 3rd place and no preference for each notation.
```{r}
# ranking prepared
ranking <- read.csv("data_prepared/data-survey-ranking_prepared.csv")
```


File: ***data-post-hoc_prepared.csv*** - contains post-hoc questions
```{r}
# read from prepared csv
posthocdata <- read.csv("data-post-hoc_prepared.csv")
posthocdata <- head(posthocdata,-1)
```



## Descriptive Statistics

### Participants

#### Profile

```{r}
# participant profiles
profileSums <- colSums(profiledata == "Yes")
profile_columns = c("Profile","Frequency")
df_p = data.frame(matrix(nrow = 0, ncol = length(profile_columns))) 
colnames(df_p) = profile_columns
df_p[nrow(df_p) + 1,] = c("Author",profileSums["author"])
df_p[nrow(df_p) + 1,] = c("User",profileSums["user"])
df_p[nrow(df_p) + 1,] = c("Researcher",profileSums["researcher"])
df_p[nrow(df_p) + 1,] = c("Other",3) # manual count for profile author:no user:no researcher:no

# convert frequency column into numeric format
df_p$Frequency = as.numeric(as.character(df_p$Frequency))

# order dataframe ascending
# this way it's displayed descending in horizontal barplot
df_p <- df_p[order(df_p$Frequency),] 

# barplot function
plot_profiles <- function() {
  counts <- table(df_p$Frequency)
  par(mar=c(3, 7, 3, 3))
  profile_plot <- barplot(height=df_p$Frequency, names=df_p$Profile, main="Participant Profiles", horiz=T,las=1,xlim=c(0,50),width = c(2,2,2,2,2))
  text(x=df_p$Frequency, profile_plot, labels = paste(df_p$Frequency), pos=4, offset=0.3, xpd=T)
}

# save barplot as EPS
setEPS()
postscript("plots/participant_profiles.eps", width = 7, height = 3.5)
plot_profiles()
dev.off()

# save barplot as PNG
png(filename="plots/participant_profiles.png")
plot_profiles()
dev.off()

# print barplot
plot_profiles()
```

#### Profile Details

```{r}
# Author 
## Years active
## How many years have you been authoring / producing educational videos, lecture
## recordings or other video-based learning content?
profiledata$authorHOWMANYYEARS <- as.numeric(as.character(profiledata$authorHOWMANYYEARS))
summary(na.omit(profiledata$authorHOWMANYYEARS))

## Videos created 
## How many educational videos, lecture recordings or other video-based learning
## content have you produced until now?
profiledata$authorHOWMANYVIDEOS <- as.numeric(as.character(profiledata$authorHOWMANYVIDEOS))
summary(na.omit(profiledata$authorHOWMANYVIDEOS))

# User 
## Years active
## How many years have you been using video-based learning content?
profiledata$userHOWMANYYEARS <- as.numeric(as.character(profiledata$userHOWMANYYEARS))
summary(na.omit(profiledata$userHOWMANYYEARS))

## Courses
## How many courses have you tought with the support of video-based learning con-
## tent? Please also count recurring courses, which possibly took place each semester.
profiledata$userHOWMANYCOURSES <- as.numeric(as.character(profiledata$userHOWMANYCOURSES))
summary(na.omit(profiledata$userHOWMANYCOURSES))

# Researcher
## Years researching
## How many years have you conducted research in the field of video-based learning
## / technology enhanced learning / educational technology?
profiledata$researchHOWMANYYEARS <- as.numeric(as.character(profiledata$researchHOWMANYYEARS))
summary(na.omit(profiledata$researchHOWMANYYEARS))

## Papers published
## How many papers have you published in the field of video-based learning? 
profiledata$researchHOWMANYPUB <- as.numeric(as.character(profiledata$researchHOWMANYPUB))
summary(na.omit(profiledata$researchHOWMANYPUB))
```



### Dependent Variables

#### Usability (SUS Score)

**Boxplot:** Visual inspection of data, using a boxplot.
```{r}
plot_boxplot_sus <- function() {
  df$notation <- factor(df$notation, levels=c("kv","cnl","sc"))
  par(mar=c(5, 5, 3, 2))
  boxplot(sus ~ notation, data = df, main = "Usability (System Usability Scale)", xlab = "Notation", ylab = "SUS score", names =c("KV","CNL","SC"))
}

# save barplot as EPS
setEPS()
postscript("plots/survey_sus_boxplots.eps")
plot_boxplot_sus()
dev.off()

# save barplot as PNG
png(filename="plots/survey_sus_boxplots.png")
plot_boxplot_sus()
dev.off()

# print boxplot
plot_boxplot_sus()
```

**Summary:** For reporting of mean (min, q1, mdn, q2, max) and sd:
```{r}
df %>%
  group_by(notation) %>%
  summarize(mean = mean(sus), sd = sd(sus), min = min(sus), max = max(sus), med = median(sus), q1 = quantile(sus, 0.25), q3 = quantile(sus, 0.75))
```

**Mean Difference:** For reporting of unstandardized mean difference:
```{r}
##### mean difference ####
# values
sus_cnl <- mean(df$sus[df$notation == "cnl"])
sus_kv <- mean(df$sus[df$notation == "kv"])
sus_sc <- mean(df$sus[df$notation == "sc"])
# difference between cnl kv: cnl-kv
sus_diffcnlkv <- sus_cnl-sus_kv
# avergae of nl kv
sus_avgcnlkv <- (sus_cnl+sus_kv)/2
sus_ratiocnlkv <- sus_diffcnlkv/sus_avgcnlkv
# percentage
sus_percent_diff_cnlkv <- abs(sus_ratiocnlkv*100)
print(c("CNL/KV mean difference in percentage: ", sus_percent_diff_cnlkv))

# difference between sc kv: sc-kv
sus_diffsckv <- sus_sc-sus_kv
# avergae of nl kv
sus_avgsckv <- (sus_sc+sus_kv)/2
sus_ratiosckv <- sus_diffsckv/sus_avgsckv
# percentage
sus_percent_diff_sckv <- abs(sus_ratiosckv*100)
print(c("SC/KV mean difference in percentage: ", sus_percent_diff_sckv))
```

**Stacked Barplots:** Detailed response results to the individual SUS questions

The following stacked bar plots have been used and reported by Ivanchikj et al. (2020) (see [References] and [DOI](https://doi.org/10.1145/3365438.3410990)). For the purpose of this thesis the style and reporting has been replicated with our data.

SUS Responses for Key-value based notation
```{r}
# In preparation for plotting we setup a function to define a limit width for strings wrapped
# Source juliasilge (comment on github from Oct 4, 2022)
# Link: https://github.com/juliasilge/tidytext/issues/222
custom_labeler <- function(x) {
  x %>%
    str_replace("___.+$", "") %>%
    str_wrap(width = 50)
}

# First we use mutate and factor to bring questions (PQ and NQ) into desired order for plotting
#kv
df_kv_tidy <- df_kv_tidy %>% mutate(question_short = substr(question, start = 1, stop = 3))

df_kv_tidy <- df_kv_tidy %>% mutate(question_short = factor(question_short, 
                                                      levels = c("PQ5",
                                                                 "PQ4",
                                                                 "PQ3",
                                                                 "PQ2",
                                                                 "PQ1",
                                                                 "NQ5",
                                                                 "NQ4",
                                                                 "NQ3",
                                                                 "NQ2",
                                                                 "NQ1")))

#cnl
df_cnl_tidy <- df_cnl_tidy %>% mutate(question_short = substr(question, start = 1, stop = 3))

df_cnl_tidy <- df_cnl_tidy %>% mutate(question_short = factor(question_short, 
                                                              levels = c("PQ5",
                                                                         "PQ4",
                                                                         "PQ3",
                                                                         "PQ2",
                                                                         "PQ1",
                                                                         "NQ5",
                                                                         "NQ4",
                                                                         "NQ3",
                                                                         "NQ2",
                                                                         "NQ1")))

#sc
df_sc_tidy <- df_sc_tidy %>% mutate(question_short = substr(question, start = 1, stop = 3))

df_sc_tidy <- df_sc_tidy %>% mutate(question_short = factor(question_short, 
                                                            levels = c("PQ5",
                                                                       "PQ4",
                                                                       "PQ3",
                                                                       "PQ2",
                                                                       "PQ1",
                                                                       "NQ5",
                                                                       "NQ4",
                                                                       "NQ3",
                                                                       "NQ2",
                                                                       "NQ1")))

# Then we factorize the Likert answer scale
#kv
df_kv_tidy <- df_kv_tidy %>% mutate(answer = factor(answer, levels = c('Strongly Disagree','Disagree','Neutral','Agree','Strongly Agree')))

#cnl
df_cnl_tidy <- df_cnl_tidy %>% mutate(answer = factor(answer, levels = c('Strongly Disagree','Disagree','Neutral','Agree','Strongly Agree')))

#sc
df_sc_tidy <- df_sc_tidy %>% mutate(answer = factor(answer, levels = c('Strongly Disagree','Disagree','Neutral','Agree','Strongly Agree')))

# The actual plots
# kv
# option 2+3: for horizontal + vertical figure use
survey_kv_sus_responses <- ggplot(df_kv_tidy, aes(x=question_short, fill=answer)) + 
  labs(title = "Key-Value Notation") +
  geom_bar(width = 0.7, position = position_stack(reverse = TRUE)) + 
  scale_fill_manual(values=c("darkred","red", "grey", "darkolivegreen1", "darkgreen")) +
  scale_y_continuous(expand = expansion(0)) +
  theme_bw() +
  theme(axis.text.y=element_text(hjust=0), axis.title.y = element_blank(), axis.title.x =element_blank(), legend.position = "bottom", legend.justification = c(1,1), legend.title = element_blank()) +
  facet_rep_grid(notion ~ ., scales = "free", repeat.tick.labels = "all") +
  scale_x_discrete(labels = custom_labeler) +
  coord_capped_cart(bottom="both", left="both") +
  coord_flip()

# cnl
# option 2+3: for horizontal + vertical figure use
survey_cnl_sus_responses <- ggplot(df_cnl_tidy, aes(x=question_short, fill=answer)) + 
  labs(title = "Controlled Natural-Language Notation") +
  geom_bar(width = 0.7, position = position_stack(reverse = TRUE)) + 
  scale_fill_manual(values=c("darkred","red", "grey", "darkolivegreen1", "darkgreen")) +
  scale_y_continuous(expand = expansion(0)) +
  theme_bw() +
  theme(axis.text.y=element_text(hjust=0), axis.title.y = element_blank(), axis.title.x =element_blank(), legend.position = "bottom", legend.justification = c(1,1), legend.title = element_blank()) +
  facet_rep_grid(notion ~ ., scales = "free", repeat.tick.labels = "all") +
  scale_x_discrete(labels = custom_labeler) +
  coord_capped_cart(bottom="both", left="both") +
  coord_flip()

# sc
# option 2+3: for horizontal + vertical figure use
survey_sc_sus_responses <- ggplot(df_sc_tidy, aes(x=question_short, fill=answer)) + 
  labs(title = "Script Notation") +
  geom_bar(width = 0.7, position = position_stack(reverse = TRUE)) + 
  scale_fill_manual(values=c("darkred","red", "grey", "darkolivegreen1", "darkgreen")) +
  scale_y_continuous(expand = expansion(0)) +
  theme_bw() +
  theme(axis.text.y=element_text(hjust=0), axis.title.y = element_blank(), axis.title.x =element_blank(), legend.position = "bottom", legend.justification = c(1,1), legend.title = element_blank()) +
  facet_rep_grid(notion ~ ., scales = "free", repeat.tick.labels = "all") +
  scale_x_discrete(labels = custom_labeler) +
  coord_capped_cart(bottom="both", left="both") +
  coord_flip()

# save
# option 3: vertical
sus_responses_all <- ggarrange(survey_kv_sus_responses, survey_cnl_sus_responses, survey_sc_sus_responses, ncol=1, nrow=3, common.legend = TRUE, legend="bottom")

bottom_text <- c("Questions: \n NQ1: I find the notation unnecessarily complex.\n NQ2: I think that I would need the support of a technical person to be able to use this notation.\n NQ3: I think there is too much inconsistency in this notation.\n NQ4: I would find the notation very cumbersome to use.\n NQ5: I would need to learn a lot of things before I could get going with this notation.\n PQ1: I think that I would like to use this notation frequently.\n PQ2: I think the notation would be easy to use.\n PQ3: I find the various functions in this notation are well integrated.\n PQ4: I would imagine that most people would learn to use this notation very quickly.\n PQ5: I would feel very confident using the notation.")

# option 3: vertical
sus_responses_all_final <- annotate_figure(sus_responses_all,
                                           top = text_grob("Individual System Usability Scale Responses", hjust = 2.3, size = 14, face = "bold"),
                                           bottom = text_grob(bottom_text,
                                                              hjust = 1, x = 1, size = 10)
)

# option 3: vertical
ggsave(file="plots/survey_all_sus_responses_vertical.svg", plot=sus_responses_all_final, width = 10, height = 20)
ggsave(file="plots/survey_all_sus_responses_vertical.eps", plot=sus_responses_all_final, width = 10, height = 20)
```
#### Ranking (Personal Preference)

**Stacked Barplot:** Visual inspection of data, using a stacked barplot.

```{r}
counts_SC <- count(ranking, RANKING.SQ001.)
counts_CNL <- count(ranking, RANKING.SQ002.)
counts_KV <- count(ranking, RANKING.SQ003.)

notation <- c(rep(c("SC"),4),rep(c("CNL"),4),rep(c("KV"),4))
ranking <- rep(c("1st place","2nd place","3rd place","no preference"), 3)  
counts <- c(counts_SC$n[2:5],counts_CNL$n[2:5],counts_KV$n[2:5])
data_ranking <- data.frame(notation,ranking,counts)

desired_order <- c("KV","CNL","SC")
# Re-order the levels
data_ranking$notation <- factor( as.character(data_ranking$notation), levels=desired_order )

# Stacked + count inside bars
survey_ranking_stacked <- ggplot(data_ranking, aes(fill=ranking, y=counts, x=notation)) + 
  geom_bar(position="fill", stat = "identity") +
  geom_text(aes(label=after_stat(counts)), stat="identity", position="fill", vjust = 1.2, size = 2) +
  scale_y_continuous(labels = scales::label_percent(scale = 55, prefix = "", suffix = "")) +
  labs(title = "Personal Ranking of Notations", subtitle = "Ranking based on personal preference for notation", x = "Notation", y = "Counts", fill = "Ranking") +
  theme_bw()

# save
ggsave(file="plots/survey_ranking_stacked.svg", plot=survey_ranking_stacked, width = 9, height = 5.87)
ggsave(file="plots/survey_ranking_stacked.eps", plot=survey_ranking_stacked, width = 9, height = 5.87)
```


### Post-hoc Questions

The following questions were asked after assessing the perceived usability for all three notations.

Preparing for likert plotting
```{r}
# get the sums of each column with likert levels
strong_disagree <- colSums(posthocdata == "1")
strong_disagree <- as.data.frame(strong_disagree)

disagree <- colSums(posthocdata == "2")
disagree <- as.data.frame(disagree)

neutral <- colSums(posthocdata == "3")
neutral <- as.data.frame(neutral)

agree <- colSums(posthocdata == "4")
agree <- as.data.frame(agree)

strong_agree <- colSums(posthocdata == "5")
strong_agree <- as.data.frame(strong_agree)
```

#### Questions 1-4

```{r, fig.width=7,fig.height=5}
# setup 
col_titles = c("Item","Strong Disagree","Disagree","Neutral","Agree","Strong Agree")
df_posthoc_all = data.frame(matrix(nrow = 0, ncol = length(col_titles))) 
colnames(df_posthoc_all) = col_titles

# fill
df_posthoc_all[1, ] <- c("Q1",strong_disagree[5,],disagree[5,],neutral[5,],agree[5,],strong_agree[5,])
df_posthoc_all[2, ] <- c("Q2",strong_disagree[6,],disagree[6,],neutral[6,],agree[6,],strong_agree[6,])
df_posthoc_all[3, ] <- c("Q3",strong_disagree[7,],disagree[7,],neutral[7,],agree[7,],strong_agree[7,])
df_posthoc_all[4, ] <- c("Q4",strong_disagree[8,],disagree[8,],neutral[8,],agree[8,],strong_agree[8,])

# convert to numeric
i <- c(2:6)  
df_posthoc_all[, i] <- apply(df_posthoc_all[, i], 2, function(x) as.numeric(as.character(x)))
#sapply(df_posthoc_all, mode)

# plot
title <- "Post-hoc Questions"
q1_plot <- plot(likert(summary = df_posthoc_all), group.order = c("Q1", "Q2", "Q3", "Q4")) +
  labs(title = title, y = "Percentage", caption = "<b>Q1 Efficiency:</b> Do you agree with the statement that a text-based approach could be more efficient?<br><b>Q2 Effectiveness:</b> Do you agree with the statement that a text-based approach could be more effective?<br><b>Q3 Textual notation as a first step:</b> Could you imagine using a textual notation as a starting point to sketch out<br>the main structure of a learning module in order to continue editing in a graphical user interface (GUI)?<br><b>Q4 Textual notation combined with GUI:</b> Could you imagine using a textual notation in combination with<br>an editing environment in a graphical user interface (GUI), allowing you to refine and switch between<br>textual notation and GUI representation?") +
  theme(plot.caption = element_markdown())
  
q1_plot

#save in variable image to export to svg
image=q1_plot
#save
ggsave(file="plots/posthoc_questions.svg", plot=image, width=7, height=5)
```

## Results and Hypothesis Tests

### Tests of Normality

#### Usability (SUS Score)

Two types of visual inspections (histograms and qq-plots) and two types of statistical tests (Shapiro-Wilk and Skewness-Test).

**Result:** Data is not normal. Boxcox transformation is used.

```{r}
# boxcox transform sus
df$sus.boxcox = (df$sus^3.2-1)/3.2
```

